---
title: "Group 1 — Clinic bodies"
subtitle: "Studio wall"
author: "Group names here"
format:
  html:
    toc: true
    toc-depth: 2
---

# Spine

> Who is well in the clinic, who is labelled ill, and where does infant death still sit on the map?

This sentence stays at the top. Every panel must attach to it. If a plot cannot be introduced in a clause that hangs off this line, it is off the wall.

# Introduction — the tables we own

- **Panel A** (`panel-A.csv`) — clinic measurements: age, blood pressure, cholesterol, maximum heart rate, plus sex and heart-disease label for grouping.
- **Panel B** (`panel-B.csv`) — the *same people*, Quals only: chest pain, blood sugar flag, exercise angina, heart disease.
- **Panel C** (`panel-C.csv`) — infant mortality by country. No geometry in the file; we join to a world map in R.

A does not grow a mosaic. B does not grow a honest histogram of chest pain. C is a rate, not a headcount.

We load from this folder:

Found tables train the mark. A campus Free Hunch, if we collect one, is a **made** panel of the same pair type — new people, old grammar. It does not replace A, B, or C.

```{r}
#| label: SETUP
library(tidyverse)
library(skimr)
library(mosaic)
library(ggformula)
```

# Panel A — body / measure

*Age, BP, Cholesterol, MaximumHR are Quant. Sex and HeartDisease are Quals used to *split* the Quants.*


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

The Qualitative variable in this table entail - HeartDisease and Sex(These were character variables which were mutated to factor variables).

the quantitative variables in this table are - Age, BP, Cholestrol, MaximumHR.

A few piecies of data missing here are the fatality rate and the fact that casuality cannot be proved since this is not an RCT data set.

```r
# glimpse() / inspect() / skim()
```

```{r}

myPanelAData <- read_csv(file = "dta/panel-A.csv")

glimpse(myPanelAData)

```

```{r}
#| label: change_to_factor
myPanelAData |> mutate(Sex = as.factor(Sex))
myPanelAData |> mutate(HeartDisease = as.factor(HeartDisease))
myPanelAMod <- myPanelAData |> dplyr::mutate(Sex = as.factor(Sex), HeartDisease = as.factor(HeartDisease))
glimpse(myPanelAMod)
```

```{r}

myPanelAData <- read_csv(file = "dta/panel-A.csv")
skim(myPanelAData)

```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

Question can answer:
1.How does cholestrol affect different sex?

2.Based on sex and age how does the maximum Heartrate fluctuate?

3.How does the age distribution of people with and without heart disease differ in this dataset?

4.how does the number of men and women having BP vary over the years?

Questions it cannot answer:

1.For what % of the population is cholestrol the reason for heart disease?

2.Does sex itself influence a person's maximum heart rate, or are the differences in maximum heart rate actually related to other factors such as age or health?

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

Questions can be answered:
```{r}
myPanelAData |> gf_boxplot(~ Cholesterol, bins = 50, fill = ~ Sex)
```

```{r}
myPanelAData |> gf_point(MaximumHR~Age, color = ~Sex) |>  gf_smooth(method = "lm") |>  gf_density_2d()
```

```{r}
myPanelAData |> gf_boxplot(Age ~ HeartDisease)
```

```{r}
myPanelAMod |> gf_line(BP ~ Age,bins= 100, color =  ~Sex)
```

Questions that cannot be answered:
```{r}
myPanelAData |> gf_point(Cholesterol~Sex,color = ~HeartDisease)
```

```{r}
myPanelAData |> ggplot(aes(Sex, MaximumHR, fill = Sex)) + geom_violin(alpha = 0.6) + geom_jitter(width = 0.08, alpha = 0.3, size = 1) + theme_minimal()
```

### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.

```{r}
myPanelAData |> group_by(Sex) |> summarize( mean_MaximumHR = mean(MaximumHR, na.rm = TRUE))
myPanelAData |> group_by(Sex) |> summarize( mean_Cholesterol = mean(Cholesterol, na.rm = TRUE))
myPanelAData |> group_by(Sex) |> summarize( mean_BP = mean(BP, na.rm = TRUE))
```

```{r}
mosaic::t_test(MaximumHR ~ Sex, data = myPanelAData)
mosaic::t_test(Cholesterol ~ Sex, data = myPanelAData)
mosaic::t_test(BP ~ Sex, data = myPanelAData)
```

# Panel B — category / talk

*ChestPain, Sex, HeartDisease, ExerciseInducedAngina. Mosaic or stacked bar. No mean of ChestPain.*


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```

```{r}
myPanelBData <- read_csv(file = "dta/panel-B.csv")

skim(myPanelBData)
```

```{r}

myPanelBData <- read_csv(file = "dta/panel-B.csv")
glimpse(myPanelBData)
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Panel C — place

*Join `country` to `rnaturalearth` / `tmap::World`. Fill is infant mortality (a rate).*

Geometry is not in the CSV. Join in R, then use the same seven headings.

```r
# library(sf)
# library(rnaturalearth)
# world <- ne_countries(scale = "medium", returnclass = "sf")
# # join on country name; list the unmatched labels
```


### 1. See the table

What is a case? Which columns are Quant / Qual? What is missing?

```r
# glimpse() / inspect() / skim()
```
```{r}
myPanelCData <- read_csv(file = "dta/panel-C.csv")
glimpse(myPanelCData)
```

```{r}

myPanelCData <- read_csv(file = "dta/panel-C.csv")
skim(myPanelCData)

```

```{r}

myPanelCData <- read_csv(file = "dta/panel-C.csv")

inspect(myPanelCData)
```

### 2. Ask

One question *this* table can answer. One it cannot. Write both.

### 3. Choose a mark

Geom, map mark, or (later) test that matches the pair type. Name it before you code.

### 4. Predict

What do we think we will see? One sentence, before Run.

### 5. Make

Human brick first. Optional: one AI prompt, pasted below, only for what section 3 named.

```r
# human chunk

# ## Prompt we actually used
# ai_draft
# ours
```

### 6. Say

Caption + one sentence we will stand behind. Title is a sentence, not a column name.

### 7. Limit

What this panel cannot do. Which kit table we refused to abuse.


# Conclusion

What the spine looks like after three panels.

- The question we will defend in ninety seconds
- What we will **not** claim
- Which lines were human, which were machine
- Limits: sample, missingness, the test we did not run
- Gift for the steal-this strip (palette, annotation, join trick)

```r
# sessionInfo()
```




